
---
title: "zoocat package"
author: "Ran-Ran He"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = FALSE,
    comment = "#>",
    fig.path = "README-"
)
```

**zoocat** package is an extension package of the **zoo** package. The aim of this package is:

* Manipulate multivariate time serie conveniently.
* Convert between time series data and data frame.
* Represent multi-dimensional data by two-dimensional format, so we can focus on one dimension.

For using **zoocat** package, first load it:

```{r, message=FALSE, highlight=TRUE}
library(zoocat)
```

# Construct a **zoocat** object

Although column names can be used to identify columns, for "`zoocat`" class, a data frame is used to store information of columns, we call it column attribute table (**cattr** table).
In the **cattr** table, each row represents a column of the core data in the object, and each column represents a column attribute. 
Using following codes to construct a **zoocat** object based on a matrix.

```{r}
mat <- matrix(round(rnorm(24), 2), ncol = 4)
ctable <- data.frame(treatment = factor(rep(c('a', 'b'), 2), levels = c('a', 'b')), 
                     site = factor(rep(c('s1', 's2'), each = 2), levels = c('s1', 's2')))
zc <- zoocat(mat, order.by = 2011 : 2014, colattr = ctable, index.name = 'year')
print(zc)
```

It can be seen that there are two column attributes: "treatment" and "site", and the index means "year".
So, each data value corresponds to three underlying variables, i.e., treatment, site and year.

# Set and get the **cattr** table

Using following codes to get and set **cattr** table.

```{r, highlight=TRUE}
cattr(zc)
zc2 <- zc
cattr(zc2) <- data.frame(cattr(zc), observer = 'Jack')
print(zc2)
```


# Merge by columns

Merging by columns is similar with "`zoo`" object except that the **cattr** tables need to be merged by rows.
Note that NA is added if it is needed.

```{r, highlight=TRUE}
zc2 <- zc + 10
cattr(zc2) <- data.frame(site = 's3', added = rep(TRUE, 4))
zc.merge <- cbind(zc, zc2)
print(zc.merge)
cattr(zc.merge)
```


# Melt and cast

Similar with package **reshape2**, you can melt and cast between data frame and **zoocat**.

```{r, highlight=TRUE}
df.melt <- melt(zc)
head(df.melt)
```

```{r, highlight=TRUE}
cast2zoocat(df.melt, index.var = 'year', value.var = 'value')
```

Note that casting from a data frame is a convenient method to build a "`zoocat`" object.


# Reset the index variable

The index variable can be reset to be a variable in the **cattr** table.

```{r, highlight=TRUE}
reset_index_var(zc, index.var = 'treatment')
```

# apply series functions

Functions can be applied for each column of the "`zoocat`" object, and the results will be binded with the **cattr** table.

```{r}
apply_col(zc, FUN = mean)
apply_col(zc, FUN = function (x) {c(mean = mean(x), sd = sd(x))})
```

Another method is to apply a function for the whole core data, and the results can be binded with the **cattr** table or index.

```{r}
apply_core(zc, FUN = colMeans, bind = 'cattr')
apply_core(zc, FUN = rowMeans, bind = 'index')
apply_core(zc, FUN = function (x) {t(x*2)}, bind = c('cattr', 'index'))
```


# Filter columns

Function `filter_col` can be used to extract columns by specified conditions on the **cattr** table:

```{r}
filter_col(zc, cond = treatment == 'a' & site == 's2')
```



# Others

For "`zoocat`" class, all methods for the "`zoo`" class can be used.
Addtionally, class "`zoomly`", based on "`zoocat`", is designed to manipulate monthly data.
 